---
title: "DAG and data issues"
output: html_notebook
---

```{r}
library(ggplot2)
```

Causality in heart disease can be understood via a Directed Acyclic Graph (DAG) which  maps the relationships between various contributing predictors. In our DAG, we focus on six  variables that we believe most significantly influence the likelihood of developing heart disease.

Ranked in order of relative impact, the most influential predictor is oldpeak, which measures ST depression on an electrocardiogram (Patil, 2025). This reflects how the heart responds under stress compared to rest, with a downward slope indicating potential dysfunction (Patil, 2025). Next is ca, which captures the number of major blood vessels showing signs of blockage (Patil, 2025). The more vessels affected, the higher the cardiovascular risk (Patil, 2025). Trestbps, recording resting blood pressure, follows closely, as high blood pressure at rest is linked to heart disease, stroke and chronic cardiovascular issues (Patil, 2025).

The fourth predictor is restecg, which looks at resting electrocardiogram results (Patil, 2025). Abnormalities here may potentially signal increased strain on the heart or underlying valvular  defects (Patil, 2025). The last two predictors, chol (cholesterol levels) and fbs (fasting blood sugar) are generally considered more manageable through lifestyle changes (Patil, 2025). Diet, exercise and medication can significantly reduce their impact, whereas the other variables often indicate deeper physiological issues that may require surgical intervention (Patil, 2025).

These predictors not only contribute individually but are also interlinked. For instance, higher fasting blood sugar can lower HDL (good cholesterol) and raise LDL (bad cholesterol) worsening lipid profiles (Patil, 2025). Greater cholesterol levels lead to plaque build-up in the arteries, narrowing them and forcing the heart to work harder (Patil, 2025). This can result in increased blood pressure, known as hypertension, which is often detected as ST depression on an ECG (Patil, 2025). Such findings might prompt further investigation through procedures like a fluoroscopy to assess the extent of cardiovascular damage (Patil, 2025).

```{r}
# DAG
heart_dag <- ggdag::dagify(chol ~ fbs,
                           trestbps ~ chol,
                           oldpeak ~ trestbps,
                           restecg ~ oldpeak,
                           ca ~ restecg,
                           target ~ fbs,
                           target ~ chol,
                           target ~ trestbps,
                           target ~ oldpeak, 
                           target ~ restecg,
                           target ~ ca)

ggdag::ggdag(heart_dag) + theme_void()
```

We encountered issues with our dataset sourced from Kaggle, namely related to counterintuitive findings that raised concerns about the accuracy of certain attributes. For instance, the dataset suggests that experiencing chest pain during exercise correlates with a lower likelihood of heart disease, while showing no chest pain increases the risk of having heart disease, an interpretation that contradicts science (Hamada, 2025). This anomaly suggests that the target variable in our Kaggle dataset may have been reversed by the uploader - where a value of 0 actually represents a diseased heart and 1 indicates a healthy heart, rather than the original convention where 0 = healthy and 1 = diseased.

Thus, while the coefficients discussed in our report remain valid in terms of magnitude and relative importance, their directional interpretation should be considered in reverse.

Take the following example, assuming the values of tha target are reversed correctly: 

If we were to include only trestbps (resting blood pressure) as a predictor in a logistic regression model against the target, the estimated effect on the log-odds of heart disease might appear stronger than it actually is. This overestimation occurs because trestbps is positively correlated with cholesterol, a variable omitted in this simplified model. When blood pressure rises, it coincides with an underlying increase in cholesterol levels. As a result, the model mistakenly attributes some of cholesterol's influence on target to trestbps, leading to a higher coefficient. This is an example of Omitted Variable Bias (OVB), where the absence of a relevant predictor, in our case cholestrol, causes the effect of an included variable to be biased in magnitude. Given the interlinked nature of these predictors, as discussed previously, it is important to interpret coefficients within a full model to avoid drawing misleading conclusions from individual variables.

Bibliography

Patil, M. (2025, April 1). Variables impacting presence of heart disease. (A. Patil, Interviewer)
Hamada, M. (2025, 16 April). Heart Disease Dataset - Discussion. Retrieved from Kaggle: https://www.kaggle.com/datasets/johnsmith88/heart-disease-dataset/discussion/171669
