---
output: html_document
---

# ST 310 Group Project: Predicting Presence of Heart Disease in Patients

**Candidate Numbers:**



## Outline/Contents




## Introduction

### The Dataset
Where from, vars used etc etc

### What We Aim to Achieve
Motivations and objectives

## Set Up

### Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# General
library(dplyr) # for filtering and sorting to be easier.
library(rstudioapi) # to set the correct working directory


# Fitting Models
#library(tidyverse)
library(ISLR) # to get all the data and models from ISLR to experiment
library(gam) # to fit gams
library(tree) # for tree fitting to guage the functional form of the data


# Plotting/ Visualisation
library(ggplot2) # for maintainable plots with a lot of interactability
library(yardstick) # for plotting an ROC curve
```

### Directories/Paths
#### Set the current directory to the working directory
```{r}
setPath <- dirname(getSourceEditorContext()$path)

setwd(setPath)
getwd()
```

#### Set the path for the data and predictions
```{r, echo=FALSE}
# Change data_path to where I saved the data
data_path <- paste0(getwd(),"/../data/")
predictions_path <- paste0(getwd(),"/../predictions/")

```

### Load Data
```{r}
Heart <- read.csv(paste0(data_path, "Heart.csv"))
# Rename the outcome column to y
Heart <- Heart %>% rename(y = target)

# Move the last column to the front so that the cross-validation works
Heart <- Heart[, c(ncol(Heart), 1:(ncol(Heart) - 1))]

# Change y to be factor so that  it is recognised as classification
Heart_non_fctr <- Heart
Heart$y <- factor(Heart$y)

head(Heart)
names(Heart)
```
### Cleaning
Any cleaning we have to do
### Exploratory Data Analysis (EDA)
Any eda and domain-specific research




## Model Fitting

### Set Up

#### Set up separate training and validation sets when k-fold cross-validation takes too long
```{r}

make_sets <- function(df, frac = 0.6, classify = F, test_bal = 0.68) {
  train_indices <- sample(1:nrow(df), size = frac * nrow(df), replace = F)
  
  y_train <- df[, 1]
  X_train <- as.matrix(df[, -1])
  
  if (frac>0) {
    valid_indices <- -train_indices
  }
  else {
    valid_indices = rownames(df)
  }
  train_fold <- df[train_indices,]
  valid_fold <- df[valid_indices,]
  
  const_yhat_success <- NA
  outcome_counts_pre <- NA
  outcome_counts_valid <- NA
  if (classify==T) {
    outcome_counts_pre <- valid_fold |> count(y)
    outcome_counts_pre$proportion <- outcome_counts_pre$n / sum(outcome_counts_pre$n)
    rare_class_size <- round(((1-test_bal) * outcome_counts_pre[outcome_counts_pre$y==1,"n"] * (1/test_bal)),0)
    rare_class_indices <- which(valid_fold$y==0)
    common_class_indices <- which(valid_fold$y==1)
    sampled_indices <- sample(rare_class_indices, size = rare_class_size, replace = FALSE)
    valid_indices <- c(sampled_indices,common_class_indices)
    valid_fold <- valid_fold[valid_indices,]
    outcome_counts_valid <- valid_fold |> count(y)
    outcome_counts_valid$proportion <- outcome_counts_valid$n / sum(outcome_counts_valid$n)
    const_yhat_success <- outcome_counts_valid$proportion[[2]]
  }
  #train_fold <- train_fold[order(rownames(train_fold)),]
  #valid_fold <- valid_fold[order(rownames(valid_fold)),]
  
  y_train_fold <- train_fold[, 1]
  X_train_fold <- as.matrix(train_fold[, -1])
  
  y_valid_fold <- valid_fold[, 1]
  X_valid_fold <- as.matrix(valid_fold[, -1])
  
  return(list(y_train = y_train, X_train=X_train, train_fold = train_fold, valid_fold = valid_fold, y_train_fold = y_train_fold, X_train_fold = X_train_fold, y_valid_fold = y_valid_fold, X_valid_fold=X_valid_fold, const_yhat_success= const_yhat_success, outcome_counts_pre = outcome_counts_pre, outcome_counts_valid = outcome_counts_valid, train_indices=train_indices, valid_indices = valid_indices))
}
make_sets_output_1 <- make_sets(df=Heart )
Heart_train_fold <- make_sets_output_1$train_fold
Heart_valid_fold <- make_sets_output_1$valid_fold

Heart_X_train_fold <- make_sets_output_1$X_train_fold
Heart_X_valid_fold <- make_sets_output_1$X_valid_fold

Heart_y_train_fold <- make_sets_output_1$y_train_fold
Heart_y_valid_fold <- make_sets_output_1$y_valid_fold
```

```{r, echo=F}
n_preds <- (dim(Heart)[2]-1)
print(paste("It seems that our data has", n_preds, "predictors"))
```
#### Look at the Class Balance in Our Data
```{r}
# We need the training set to be balanced, yet the validation set to have 68% balance.
outcome_counts_fold <- Heart_valid_fold |> count(y)
outcome_counts_fold$proportion <- outcome_counts_fold$n / sum(outcome_counts_fold$n)
const_yhat_success <- outcome_counts_fold$proportion[[2]]
outcome_counts_fold
```

Nice, it seems rather balanced!
#### Create A Dataframe to Store All of Our Model Types and Misclass Estimates
```{r}
gen_models_df <- function(len=10, accuracy_measures = c("min_mse")) {
  df = data.frame(
    model_type = character(len),
    model_name = character(len)
  )
  for (measure in accuracy_measures) {
    df[[measure]] <- numeric(len)
  }
  #tp stands for tuning parameter
  df$tp1_name = character(len)
  df$tp2_name = character(len)
  df$tp3_name = character(len)
  df$tp1_value = numeric(len)
  df$tp2_value = numeric(len)
  df$tp3_value = numeric(len)

  return(df)
}
Heart_models <- gen_models_df(accuracy_measures = c("cv_misclass_rate"))

Heart_models

```
#### Define a function that helps to validarte or cross-validate a grid of tuning parameters
```{r }
validate <- function(train_df = Heart, method = "cv", model_f = "tree", grid_df = tree_grid, k=5, frac = 0.6, matrix = F, classify=F, test_bal = 0.68, error_type = "mse") {
  model.f <- get(model_f)
  n_tps <- ncol(grid_df)
  folds <- sample(rep(1:k, length.out = nrow(train_df)))
  sets_output <- make_sets(df = train_df, frac=frac, classify=classify, test_bal=test_bal)
  X <- sets_output$X_train
  Y <- sets_output$y_train
  train_indices <- sets_output$train_indices
  valid_indices <- sets_output$valid_indices
  fits <- c()
  preds <- list()
  valids <- list()
  for (i in 1:nrow(grid_df)) {
    e_folds <- c()
    tree_sizes <- c()
    if (matrix==T) {
      model_input <- list()
      if (model_f=="gbart") {
        xname = "x.train"
        yname = "y.train"
      }
      else {
        xname = "x"
        yname = "y"
      }
      model_input[[xname]] <- X
      model_input[[yname]] <- Y
    }
    
    else {
      model_input <- list(
      formula = as.formula("y ~ ."), 
      data = train_df
      )
    }
    for (tp in names(grid_df)[1:n_tps]) {
      if (all(is.na(grid_df[[tp]]))==F) {
        if (tp=="mindev") {
          model_input[["control"]] <- tree.control(nobs = length(train_df$y), mindev = grid_df[i,tp])
        }
        else if (tp=="preds_used") {
          model_input[["formula"]] <- as.formula(paste("y ~ ", grid_df[i,tp]))
        }
        else if (tp!="thresh") {
            model_input[[tp]] <- grid_df[i,tp]
        }
      }
    }
  if (method == "cv") {
    for (fold in 1:k) {
        X_train_fold <- X[folds != fold, ]
        y_train_fold <- Y[folds != fold]
        X_valid_fold <- X[folds == fold, ]
        y_valid_fold <- Y[folds == fold]
        temp_df <- data.frame(y = y_valid_fold, X_valid_fold)
        sets_output <- make_sets(df = temp_df, frac=0, classify=classify, test_bal=test_bal)
        X_valid_fold <- sets_output$X_valid_fold
        y_valid_fold <- sets_output$y_valid_fold
        valid_df <- data.frame(y_valid_fold, X_valid_fold)
      if (matrix==T) {
        model_input[[xname]] <- X_train_fold
        model_input[[yname]] <- y_train_fold
        model <- do.call(model.f, model_input)
        predictions <- predict(model, X_valid_fold)
      }
      else {
        train_fold <- train_df[folds!=fold,]
        valid_fold <- valid_df
        model_input$data <- train_fold
        model <- do.call(model.f, model_input)
        if (classify==T) {
          prob_predictions <- predict(model, newdata = valid_fold, type = "response")
          predictions <- as.factor(as.numeric(prob_predictions > grid_df[i,"thresh"]))
        }
        else {
          predictions <- predict(model, valid_fold)
        }
      } 
      
      fits <- c(fits, model)
      preds[[paste0(i,"_",k)]] <- predictions
      valids[[paste0(i,"_",k)]] <- y_valid_fold
      if (classify==T ) {
        e_fold <- sum(predictions != (as.numeric(y_valid_fold)-1)) / length(y_valid_fold)
        e_folds <- c(e_folds, e_fold)
      }
      else {
        e_fold <- mean((predictions - y_valid_fold)^2)
        e_folds <- c(e_folds, e_fold)
      }
      
      
      if (model_f == "tree") tree_sizes <- c(tree_sizes, length(unique(model$where)))
    }
  }
  else if (method=="v") {
    X_train_fold <- X[train_indices, ]
    y_train_fold <- Y[train_indices]
    X_valid_fold <- X[valid_indices, ]
    y_valid_fold <- Y[valid_indices]
    if (matrix==T) {
        model_input[[xname]] <- X_train_fold
        model_input[[yname]] <- y_train_fold
        model <- do.call(model.f, model_input)
        predictions <- predict(model, X_valid_fold)
        if (model_f=="gbart") {
          predictions <- colMeans(predictions)
        }
    }
    else {
        train_fold <- train_df[train_indices,]
        valid_fold <- train_df[valid_indices,]
        model_input$data <- train_fold
        model <- do.call(model.f, model_input)
        if (classify==T) {
          prob_predictions <- predict(model, newdata = valid_fold, type = "response")
          predictions <- as.factor(as.numeric(prob_predictions > grid_df[i,"thresh"]))
        }
        else {
          predictions <- predict(model, valid_fold)
        }
        
    }
    fits <- c(fits, model)
    preds[[i]] <- predictions
    valids[[i]] <- y_valid_fold
    if (classify==T ) {
        e_fold <- sum(predictions != (as.numeric(y_valid_fold)-1)) / length(y_valid_fold)
        e_folds <- c(e_folds, e_fold)
    }
    else {
        e_fold <- mean((predictions - y_valid_fold)^2)
        e_folds <- c(e_folds, e_fold)
    }
    
    
    if (model_f == "tree") tree_sizes <- c(tree_sizes, length(unique(model$where)))
  }
  
  grid_df[[paste0(method,"_", error_type)]][i] <- mean(e_folds)
  if (model_f == "tree") grid_df$tree_size[i] <- mean(tree_sizes)

  }
return(list(grid_df = grid_df, fits = fits , preds=preds , valids = valids, train_indices=train_indices, method=method, model_f = model_f))
}
```



### Baseline Model: Logistic (Few Predictors)
Explain why we used this

#### Validate/Cross-Validate Best Tuning Parameters
Run some code
#### Analyse Results
Look at results from best Model



### Gradient-Descent-Based Model: Logistic (Few Predictors)
Take a decreasing size step, and use a loss function without a constant
Explain why we used this
#### Explain how this works (Stickiest Part of Project)

#### Validate/Cross-Validate Best Tuning Parameters
Run some code
#### Analyse Results
Look at results from best Model


### Relatively Interpretable Model: Logistic (Multiple Predictors and Interactions)

#### Validate/Cross-Validate Best Tuning Parameters
Run some code
#### Analyse Results
Look at results from best Model



### High-Dimensional Model: Regularised Logistic (Few Predictors)

#### Validate/Cross-Validate Best Tuning Parameters
Run some code
#### Analyse Results
Look at results from best Model



### Predictively Accurate Model: Classification Random Forest

#### Validate/Cross-Validate Best Tuning Parameters


**Start with Classification Tree**
```{r, eval=F}
mindev_values <- c(5,4,3,2,1)*rep(10^seq(-2, -4, by = -1), each=5) # i've done too much matrix stuff in my finance module so this dodgy syntax is now all too natural.

tree_grid <- expand.grid(mindev = mindev_values)
#grid_results$tree_size <- NA
#tree_grid <- data.frame(dummy = rep(NA,4))
tree_grid
```
```{r, eval=F} 
tree_df_output <- validate(method = "cv")
tree_df <- tree_df_output$grid_df
tree_df

```

```{r}
# Fit a tree
tree_fit <- tree(y ~ ., Heart_train_fold)
# Fit a glm
heart_2 <- glm(y~ ., Heart_non_fctr, family = "gaussian")
# Store the Probability predictions (the second row corresponds to the postive class probability)
tree_fit_prob_preds <- predict(tree_fit, Heart_valid_fold)[,2]

#GHet the standard bayesian classifier class predictions
preds <- predict(tree_fit, Heart_valid_fold, type = "class")
plot(preds, as.numeric(Heart_valid_fold$y)-1)

```

#### Cross-Validate the Best Pruning Coefficient, K (Penalty for number of terminal nodes) and overall tree size.
```{r}
cv.tree_fit <- cv.tree(tree_fit, FUN = prune.misclass)
par(mfrow = c(1, 2))
plot(cv.tree_fit$size, cv.tree_fit$dev, type = "b")
plot(cv.tree_fit$k, cv.tree_fit$dev, type = "b")
```
#### Analyse the confusion matrices for the tree model.
```{r}

tree_fit_predictions <-data.frame(
  y = Heart_valid_fold$y,
  .fitted = tree_fit_prob_preds
) |> 
  mutate(yhat = factor(as.numeric(.fitted > const_yhat_success))) |>
  mutate(misclass = (y!=yhat))
paste("Class Balance:")
tree_fit_predictions |> pull(.fitted) |> mean() # same as mean(tree_fit_predictions$.fitted)
paste("Misclass Rate:")
sum((tree_fit_predictions$y!=tree_fit_predictions$yhat)) / nrow(tree_fit_predictions)
```
**Look at ROC Curve to guage Trade-Off**
```{r, eval=F}
tree_fit_predictions |>
  roc_curve(truth = y, .fitted,
            event_level = "second") |>
  autoplot()
# the 'truth' is just the training y here, not the testing y, which we do not have, nor a validation set.
```
**Classify at different thresholds.**

```{r, eval=F}
higher_cutoff <- const_yhat_success + .15
confusion_matrix_higher <-
  tree_fit_predictions |>
  mutate(yhat = factor(as.numeric(.fitted > higher_cutoff))) |> # this line predicts y = 1 ONLY if the predicted conditional probability is above the higher cutoff
  conf_mat(truth = y, estimate = yhat)
```
```{r, eval=F}
confusion_matrix_medium <-
  tree_fit_predictions |>
  mutate(yhat = factor(as.numeric(.fitted > const_yhat_success))) |> 
  conf_mat(truth = y, estimate = yhat)
```


```{r, eval=F}
lower_cutoff <- const_yhat_success - 0.15
confusion_matrix_lower <-
  tree_fit_predictions |>
  mutate(yhat = factor(as.numeric(.fitted > lower_cutoff))) |>
  conf_mat(truth = y, estimate = yhat)
```

**1. Balanced Confusion Matrix**
```{r, eval=F}
confusion_matrix_medium |> autoplot(type = "heatmap")
```
**3. Low Cut-Off Confusion Matrix**
```{r, eval=F}
confusion_matrix_lower |> autoplot(type = "heatmap")
```
**3. High Cut-Off Confusion Matrix**
```{r, eval=F}
confusion_matrix_higher |> autoplot(type = "heatmap")
```
**Comparing key sum stats:**

```{r, eval=F}
higher_summary <- summary(confusion_matrix_higher) |>
  mutate(higher = .estimate) |>
  dplyr::select(.metric, higher)
medium_summary <- summary(confusion_matrix_medium) |>
  mutate(medium = .estimate) |>
  dplyr::select(medium)
lower_summary <- summary(confusion_matrix_lower) |>
  mutate(lower = .estimate) |>
  dplyr::select(lower)
cbind(higher_summary, medium_summary, lower_summary) |>
  knitr::kable()


```

As seen from the confusion matrices earlier, accuracy benefits the lower and not the higher threshold.

#### Analyse Results
Look at results from best Model


## Conclusion

Did we achieve our objective? Why/Why not?

## Bibliography


## Annex
Only if you get disgusting enough to go really technical. OR, if we tried something first, and it didn't wuite work, we can show that here.