---
output: html_document
---

# ST 310 Group Project: Predicting Presence of Heart Disease in Patients

**Candidate Numbers:**

```{r}
library(ISLR)
print("hello folks")
```

## Outline/Contents




## Introduction

### The Dataset
Where from, vars used etc etc

### What We Aim to Achieve
Motivations and objectives

## Set Up

### Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(tidyverse)
library(gam) # to fit gams
library(rstudioapi) # to set the correct working directory
library(tree) # for tree fitting to guage the functional form of the data
```


```{r}
setPath <- dirname(getSourceEditorContext()$path)

setwd(setPath)
getwd()
```
### Cleaning
Any cleaning we have to do
### Exploratory Data Analysis (EDA)
Any eda and domain-specific research




## Model Fitting


### Baseline Model: Logistic (Few Predictors)
Explain why we used this

#### Validate/Cross-Validate Best Tuning Parameters
Run some code
#### Analyse Results
Look at results from best Model



### Gradient-Descent-Based Model: Logistic (Few Predictors)
Take a decreasing size step, and use a loss function without a constant
Explain why we used this
#### Explain how this works (Stickiest Part of Project)

#### Validate/Cross-Validate Best Tuning Parameters
Run some code
#### Analyse Results
Look at results from best Model


### Relatively Interpretable Model: Logistic (Multiple Predictors and Interactions)

#### Validate/Cross-Validate Best Tuning Parameters
Run some code
#### Analyse Results
Look at results from best Model



### High-Dimensional Model: Regularised Logistic (Few Predictors)

#### Validate/Cross-Validate Best Tuning Parameters
Run some code
#### Analyse Results
Look at results from best Model



### Predictively Accurate Model: Classification Random Forest

#### Validate/Cross-Validate Best Tuning Parameters
Run some code
#### Analyse Results
Look at results from best Model


## Conclusion

Did we achieve our objective? Why/Why not?

## Bibliography


## Annex
Only if you get disgusting enough to go really technical. OR, if we tried something first, and it didn't wuite work, we can show that here.